---
title: "LGCP Queensland"
author: "Anthony Timings, ID 663769672"
date: "09/08/2020"
output: html_document
---

```{r setup, include=FALSE,message=FALSE,results='asis'}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,cache = TRUE, message = FALSE,warning=FALSE)
library(kableExtra)
```

```{r}
library(stelfi)
library(sf)
library(raster)
library(sp)
library(raster)
library(rgdal)

```

# Data & tidying

```{r}
ausdata <- read.csv("ausdata.csv")
```

```{r}
aus <- raster::getData('GADM', country='AUS', level=1)
class(aus)
aus <- spTransform(aus, "+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
")
```



```{r}
area_aus <- sum(raster::area(aus)) ## (m)
area_auskm2 <- area_aus/1000^2 
area_auskm2
spatial_murder_rate_aus <- nrow(ausdata)/area_auskm2 
temporal_murder_rate_aus <- nrow(ausdata)/length(table(ausdata$Year)) 
st_murder_rate <- (nrow(ausdata)/area_auskm2)/length(table(ausdata$Year)) 
```

Transform `data.frame` to `SpatialPointsDataFrame`

(EPSG:3112)
https://gis.stackexchange.com/questions/268643/converting-australian-coordinates-epsg3112-for-mapping-in-r

```{r sp}
murders_aus_sp <- ausdata
coordinates(murders_aus_sp) <- c("Longitude","Latitude")
proj4string(murders_aus_sp) <- CRS("+proj=longlat +datum=WGS84")
murders_aus_sp <-  spTransform(murders_aus_sp, 
                           CRS("+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
"))
```


There are some points located in islands outside of Australia - we'll remove these

```{r}
x = c(rep(-1212000,2), rep(3266000,2))
y = c(500000, 4553000, 4553000, 500000)
coords <- data.frame(x,y)
coordinates(coords) <- c("x","y")
proj4string(coords) <-  CRS("+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
")
## create spatial polygon(s)
poly <- SpatialPolygons(list(Polygons(list(Polygon(coords)), ID="New Aus")),
                        proj4string = CRS("+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
"))
ind = over(murders_aus_sp, poly)
ind[is.na(ind)] = 0
ind = as.logical(ind)
murders_aus_sp = murders_aus_sp[ind,]
murders_aus_sp
```

```{r plot,results='asis',fig.width=10,fig.height=10,echo = FALSE,warning=FALSE,message=FALSE}
plot(murders_aus_sp, col="red")
plot(aus, add=T)
```

Homicides are clearly concentrated in specific parts of the country.


# Queensland 

```{r}
aus2 <- raster::getData('GADM', country='AUS', level=2)
qsld <- subset(aus2, aus2$NAME_1 == "Queensland") 
qsld <- spTransform(qsld, "+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
")
```

```{r}
ausdata <- ausdata[ind,]
qsld_ind <- over(murders_aus_sp, qsld)
qsld_data <-  ausdata[which(!is.na(qsld_ind$NAME_0)),]
nrow(qsld_data)
```

We have 448 homicides in Queensland (over our entire time span).


```{r}
murders_qsld_sp <- qsld_data
head(murders_qsld_sp)
## project longitude & latitude to NZTMs
coordinates(murders_qsld_sp) <- c("Longitude","Latitude")
proj4string(murders_qsld_sp) <- CRS("+proj=longlat +datum=WGS84")
murders_qsld_sp <-  spTransform(murders_qsld_sp, 
                           CRS("+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
"))
```

```{r,results='asis',fig.width=10,fig.height=10,echo = FALSE,warning=FALSE,message=FALSE}
plot(murders_qsld_sp, col="red")
plot(qsld, add=T)
```

We want to subset the Queensland shape file to only inlcude the South-East part (includes Brisbane, Gold Coast, Sunishine Coast etc.)

```{r}
table(over(murders_qsld_sp, qsld)$NAME_2)
```

```{r, include = F}
library(sf)
library(raster)
library(sp)
a <- st_as_sf(qsld)
qsld_proj <- st_transform(a, 4326)

library(leaflet)
leaflet(data = qsld_proj) %>%
  addTiles() %>%
  addPolygons(fillOpacity = 0.8, 
                weight = 1,
                popup = qsld_proj$NAME_2)
```



```{r,results='asis',fig.width=10,fig.height=10,echo = FALSE,warning=FALSE,message=FALSE}
plot(murders_qsld_sp, col="red")
plot(qsld[qsld$NAME_2 %in% c("Gold Coast", "Brisbane",  "Sunshine Coast", "Moreton Bay", "Redland", "Logan", "Ipswich", "Somerset", "Gympie"),], add=T)
```

We capture 227 murders with this small subset. Call this bris (Brisbane). 

```{r}
bris <- qsld[qsld$NAME_2 %in% c("Gold Coast", "Brisbane",  "Sunshine Coast", "Moreton Bay", "Redland", "Logan", "Ipswich"),]
#have excluded Gympie
bris_murders= over(murders_qsld_sp, bris)
w = which(!is.na(bris_murders$NAME_0))
bris_murders = murders_qsld_sp[w,]
#we have 227 murders in this area

plot(bris)
plot(bris_murders, col="red", add=T)
```


# Covariates 

Will use Statistical Area 1 geographies for covariates. Most SA1s have a population of between 200 to 800 persons with an average population of approximately 400 persons, so covariate values should be reflective of local area at each point.

For data: https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/2033.0.55.0012016?OpenDocument

```{r}
aus_sa1 <- readOGR("sa1_2016_aust_shape/SA1_2016_AUST.shp", layer = ogrListLayers("sa1_2016_aust_shape/SA1_2016_AUST.shp"))
aus_sa1$SA1_MAIN16 = as.numeric(aus_sa1$SA1_MAIN16)
```

Read in csv with deprivation, education scores etc.
```{r}
aus_sa1_dep <- read.csv("AusDepData.csv")
aus_sa1_merged <- merge(aus_sa1, aus_sa1_dep, by.x= "SA1_MAIN16", by.y="X2016SA1Code")
```

Transform to same coordinate system as AUS. Then do over SA1 and then use this shapefile subset.

```{r}
aus_sa1 <- spTransform(aus_sa1_merged, "+proj=lcc +lat_1=-28 +lat_2=-36 +lat_0=-32 +lon_0=135 +x_0=1000000 +y_0=2000000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
")

aa = over(aus_sa1, bris)
bris_sa1 <- aus_sa1[which(!is.na(aa$NAME_0)),]
```

# Mesh

```{r Create mesh}
mesh <- inla.mesh.2d(loc.domain = broom::tidy(bris_sa1)[,1:2],
                     max.edge = c(20000,30000), cutoff = 7000)
plot(mesh)
```

# Model - spatial only

```{r, warning = F, eval=F}
bris_fit1 <-fit_lgcp_inla(mesh = mesh, locs = coordinates(bris_murders), sp = bris_sa1,
                     return.attributes = TRUE)
bris_fit1$summary.fixed
```


```{r, eval=F}
## expected number of murders at each mesh node
## areas/weights at each mesh nodes, only returned if return.attribures = TRUE
weights <- attributes(bris_fit1)$weights
ins <- which(weights != 0)
en <- exp(as.numeric(bris_fit1$summary.fixed[1]))*weights[ins]
sum(en) ## expected number across area, observed 448
```

Estimated mean of the assumed Gaussian Markov Random Field

```{r,message= F, warning=F, fig.width=10,fig.height=10}
fields <- stelfi::get_fields(bris_fit1, mesh, mean = TRUE)
grfs <- fields[[1]]
show_field(grfs, mesh, dims = c(300,300),
         col = RColorBrewer::brewer.pal(5, "Blues"), sp = bris_sa1,
             rast = FALSE)
plot(bris_murders, add = TRUE, cex = 0.2)

```

# Adding a covariate

```{r Create covariates}
bris_sa1$EcoDisScaled = scale(as.numeric(bris_sa1$EcoAdvDisScore))
bris_sa1$EducScaled = scale(as.numeric(bris_sa1$EducScore))
bris_sa1$PopDens = bris_sa1$UR.Population/bris_sa1$AREASQKM16
bris_sa1$PopDensScaled = scale(bris_sa1$PopDens)
```


Covariates include Economic Disadvantage, Education and Population Density. For Education and Economic disadvantage we use scores (mean 1000) and then scale them.

We will add population density as a covariate as an example. Covariates are defined at the mesh nodes (integration points inside the study region) and at the point locations. 


```{r, eval=F}
popdens_mesh = sp::over(SpatialPoints(mesh$loc[,1:2], proj4string = CRS(proj4string(bris_murders))), aus_sa1)$PopDensScaled
popdens_obs = sp::over(SpatialPoints(mesh$loc[,1:2], proj4string = CRS(proj4string(bris_murders))), aus_sa1)$PopDensScaled

covs = data.frame(popdens = c(popdens_mesh, popdens_obs))

```



```{r, eval = F}
bris_fit2 <-fit_lgcp_inla(mesh = mesh, locs = coordinates(bris_murders), covariates = covs, sp = bris_sa1, return.attributes = TRUE)
bris_fit2$summary.fixed
```



Estimated mean of the assumed Gaussian Markov Random Field

```{r,message= F, warning=F, fig.width=10,fig.height=10, eval=F}
fields <- stelfi::get_fields(bris_fit2, mesh, mean = TRUE)
grfs <- fields[[1]]
show_field(grfs, mesh, dims = c(300,300),
         col = RColorBrewer::brewer.pal(5, "Blues"), sp = bris_sa1,
             rast = FALSE)
plot(bris_murders, add = TRUE, cex = 0.2)

```










